Concurrent Programming with CXL

### 2020-06-14 book/paper.tex
Latest

\chapter{Monitors}

@@

>>>>

\chapter{Monitors}
\label{ch:monitors}

@@

To illustrate how Mesa condition variables are used in practice, we demonstrate, once again,
using an implementation of reader/writer locks.
Figure~\ref{fig:RWcv} shows the code.  \texttt{rwlock} is the global lock or mutex for the
reader/writer lock.  There are two condition variables: readers wait on \texttt{rcond} and
writers wait on \texttt{wcond}.  The implementation also keeps track of the number of
readers and writers in the critical section.

When inspecting the code, one important feature to notice is that \texttt{wait} is always
invoked within a \texttt{while} loop that checks for the condition that the process is
waiting for.  We explained why in Chapter~\ref{ch:monitors}: when a process resumes after
being notified, it is not guaranteed that the condition it was waiting for holds, even
if it did at the time of invoking \texttt{notify} or \texttt{notifyAll}.
It is \emph{imperative} that you always have a \texttt{while} loop around any invocation
of \texttt{wait}.

In \texttt{release\_rlock}, notice that \texttt{notify \&(wcond)} is invoked when there
are no readers left, \emph{without} checking if there are writers waiting to enter.
With Mesa monitors this is ok, because calling \texttt{notify} is a no-op if no process
is waiting.

\texttt{release\_wlock} executes \texttt{notifyAll \&(wcond)} as well as
\texttt{notify \&(wcond)}.
Again, because we do not keep track of the number of waiting readers or writers, we
have to conservatively assume that all waiting readers can enter, or, alternatively,
up to one waiting writer can enter.  So \texttt{release\_wlock} wakes up all
potential candidates.
There are two things to note here.  First, unlike split binary semaphores or Hoare
monitors, where multiple waiting readers would have to be signaled one at a time in a
baton-passing fashion (see Figure~\ref{fig:splitsema}), with Mesa monitors
all readers are awaked in one fell swoop using \texttt{notifyAll}.
Second, both readers and writers are awakened---this is ok because both execute
\texttt{wait} within a \texttt{while} loop, re-checking the condition that they
are waiting for.  So if both type of processes are waiting, either all the readers
get to enter next or one of the writers gets to enter next.
